Hybrid MILP to efficiently and accuratly \\ solve hard DNN verification instances

### 2024-09-16 ICLR25/Paper_noProof.tex
table updated in main document

\title{Partial MILP to improve $\alpha,\beta$-CROWN accuracy \\ for hard DNN verification instances}

@@

\begin{abstract}
$\alpha,\beta$-CROWN has won the last 4 VNNcomp(etitions), as the DNN verifier with the best 
trade-off between accuracy vs computational time. VNNcomp however is focusing on relatively easy verification instances (network, inputs (images)), 
with few {\em unstable nodes}. In this paper, we consider harder verification instances, namely the usual local robustness problem, and 
compare how different solutions out there fare. On such instances, $\alpha,\beta$-CROWN displays a large gap ($20-58\%$) between instances that can be verified, and instances with an explicit attack. Enabling much larger time-outs for $\alpha,\beta$-CROWN only improves verification rate by few percents, leaving a large gap of undecided instances while already taking a considerable amount of time. Resorting to other techniques, such as complete verifiers, does not fare better even with very large time-outs (10000s per instance): They would theoretically be able to close the gap, but with an untractable runtime on all but small instances, because the instances considered are hard.

In this paper, we study the complete MILP encoding of ReLU-DNNs, provide new deep insights in the LP relaxation, which allows us to carefully craft a {\em partial MILP} solution which only considers few neurons as integer variables, the rest using the LP relaxation. Compared with previous attempts, we can reduce the number of integer variables by around 4 times while maintaining the same level of accuracy. This gives rise to a very efficient yet accurate algorithm. Implemented as $\alpha,\beta$-pMILP, calling first $\alpha,\beta$-CROWN with a short time-out (10s) to solve easier instances, and then partial MILP for those for which $\alpha,\beta$-CROWN fails, produces a very accurate yet efficient verifier, reducing tremendously the gap of undecided instances to $\approx 10\%$, while keeping a reasonable runtime ($<500s$ on average per instance).
\end{abstract}

@@

\begin{tabular}{||c|c|c||c|c|c||}
	\hline \hline
	Network & Acc & Upper  & $\alpha,\beta$-Crown& $\alpha,\beta$-Crown & $\alpha,\beta$-Crown \\ 
	Perturbation &   & Bound & TO=10s & TO=30s & TO=2000s\\ \hline
	MNIST 6$\times$100 & 96\% & 90\% & 33\% & 35\% & 40\%   \\
	$\epsilon = 0.026$ &  &  & ? &  18s &  1600s  \\  \hline
    MNIST 6$\times$200 & 97\%  & 96\%  & 46\%  & 49\%  & 50\%   \\ 
	$\epsilon = 0.015$ & &  & ? &  17s &  ?s  \\  \hline
	MNIST 9$\times$100 & 95\%  & 86\%  & 23\%  & 28\%  & 28\%   \\
	$\epsilon = 0.026$ &  &  & ? &  20s &  ?s  \\  \hline
	MNIST 9$\times$200 & 95\%  & 91\%  & 35\%  & 36\%  & 37\%   \\ 
	$\epsilon = 0.015$ & &  & 7s &  s &  ?s  \\  \hline
	MNIST 6$\times$500 &  ? & 94\%  & 41\%  & 43\%  & 44\%   \\ 
	$\epsilon = 0.035$ & &  & 6s &  15s &  1002s  \\  \hline
	CIFAR CNN-B-adv & ?  & 62\%  &  40\% & 41\%  & 42\%   \\
	$\epsilon = 2/255$ &  &  & 15s & 32s & 373s  \\ \hline \hline
	CIFAR ResNet & 29\%  & 25\%  & 25\%  & 25\%  & 25\%   \\
	$\epsilon=2/255$ & : &  & 2s & 2s & 2s  \\ \hline \hline

\end{tabular}

@@

\begin{tabular}{||c||c|c|c||}
	\hline
	Network &  Marabou  & Full MILP & NNenum? \\ \hline
	MNIST 6$\times$100 & 26\% & 30\% & ?\%   \\ \hline
	$\epsilon = 0.026$ & 6520s & 7000s s
	 &  ?s  \\  \hline
\end{tabular}

@@

Gap between verified and attackable images. Lower is better.
%The table of undecided images among first 100 images in database by different methods for different network.

\begin{tabular}{|c|c|c|c|c|c|}
	\hline
	Network & $\alpha,\beta$-Crown, 10s & $\alpha,\beta$-Crown, 30s & $\alpha,\beta$-Crown, 2000s & $\alpha,\beta$-Crown-refine & Our method \\ \hline
	& Gap(\%) \hfill Time(s) & Gap(\%) \hfill Time(s) & Gap(\%) \hfill Time(s) & Gap(\%) \hfill Time(s) & Gap(\%) \hfill Time(s)\\ \hline
	MNIST 6$\times$100 & 57 \hfill 6.9 & 55 \hfill 18.9 & 50 \hfill 1026.1 & 13 \hfill 91.7 & Rate \hfill Time \\ \hline
	MNIST 6$\times$200 & 50 \hfill 6.5 & 47 \hfill 16.6 & 46 \hfill 930.3 & 9 \hfill 79.8 & Rate \hfill Time \\ \hline
	MNIST 9$\times$100 & 63 \hfill 7.2 & 58 \hfill 20.1 & 58 \hfill 1162.9 & 21 \hfill 102.2 & Rate \hfill Time \\ \hline
	MNIST 9$\times$200 & 56 \hfill 6.8 & 55 \hfill 18.2 & 54 \hfill 1083.1 & 16 \hfill 83.2 & Rate \hfill Time \\ \hline
	0 & 0 & 0 & 0 & 0 & 0 \\ \hline
\end{tabular}

>>>>

\title{Hybrid MILP to efficiently and accuratly \\  solve hard DNN verification instances}

@@

\begin{abstract}
$\alpha,\beta$-CROWN has won the last 4 VNNcomp(etitions), as the DNN verifier with the best 
trade-off between accuracy vs computational time. VNNcomp however is focusing on relatively easy verification instances (network, inputs (images)), 
with few {\em unstable nodes}. In this paper, we consider harder verification instances, namely the usual local robustness problem. On such instances, $\alpha,\beta$-CROWN displays a large gap ($20-58\%$) between instances that can be verified, and instances with an explicit attack. Enabling much larger time-outs for $\alpha,\beta$-CROWN only improves verification rate by few percents, leaving a large gap of undecided instances while already taking a considerable amount of time. Resorting to other techniques, such as complete verifiers, does not fare better even with very large time-outs: They would theoretically be able to close the gap, but with an untractable runtime on all but small {\em hard} instances.

In this paper, we study the MILP encoding of ReLU-DNNs, provide new deep insights in the LP relaxation, which allows us to carefully craft a {\em partial MILP} solution which only considers few neurons as integer variables, the rest using the LP relaxation. Compared with previous attempts, we can reduce the number of integer variables by around 4 times while maintaining the same level of accuracy. This gives rise to a very efficient yet accurate algorithm. Implemented in {\em Hybrid MILP (hMILP)}, calling first $\alpha,\beta$-CROWN with a short time-out to solve easier instances, and then partial MILP for those for which $\alpha,\beta$-CROWN fails, produces a very accurate yet efficient verifier, reducing tremendously the gap of undecided instances to $\approx 10\%$, while keeping a reasonable runtime ($<500s$ on average per instance).
\end{abstract}

@@

\begin{table}
	\centering
	\begin{tabular}{||l|c|c||c|c|c||}
		\hline
		Network & Acc & Upper  & $\alpha,\beta$-Crown& $\alpha,\beta$-Crown & $\alpha,\beta$-Crown \\ 
		Perturbation &   & Bound & TO=10s & TO=30s & TO=2000s\\ \hline
		MNIST 6$\times$100 & 99\% & 90\% & 33\% & 35\% & 40\%   \\
		$\epsilon = 0.026$ &  &  & 6.9s &  18.9s &  1026.1s  \\  \hline
		MNIST 6$\times$200 & 99\%  & 96\%  & 46\%  & 49\%  & 50\%   \\ 
		$\epsilon = 0.015$ & &  & 6.5s &  16.6s &  930.3s  \\  \hline
		MNIST 9$\times$100 & 97\%  & 86\%  & 23\%  & 28\%  & 28\%   \\
		$\epsilon = 0.026$ &  &  & 7.2s &  20.1s &  930.3s  \\  \hline
		MNIST 9$\times$200 & 97\%  & 91\%  & 35\%  & 36\%  & 37\%   \\ 
		$\epsilon = 0.015$ & &  & 6.8s &  18.2s &  1083.1s  \\  \hline
		MNIST 6$\times$500 & 100\%  & 94\%  & 41\%  & 43\%  & 44\%   \\ 
		$\epsilon = 0.035$ & &  & 6.4s &  16.4s &  1002.5s  \\  \hline
		CIFAR CNN-B-adv & 78\%  & 62\%  &  34\% & 40\%  & 42\%   \\
		$\epsilon = 2/255$&  &  & 4.3s & 8.7s & 373s  \\ \hline \hline
		CIFAR ResNet & 29\%  & 25\%  & 25\%  & 25\%  & 25\%   \\
		$\epsilon = 2/255$ &  &  & 2s & 2s & 2s  \\ \hline
	\end{tabular}
	\caption{Verification $\%$ of $\alpha,\beta$-Crown run on 7 DNNs (average runtime per image in second). The 6 first DNNs are hard instances, while the last one (ResNet) is an easy instance (it is trained to be easy to verify using Wong, but with a very low accuracy level), provided for reference.}
	\label{table_beta}
\end{table}

@@

\begin{table}
	\centering
\begin{tabular}{||l|c|c||c|c|c||}
	\hline
	Network &  Acc & Upper Bound & Marabou 2.0  & Full MILP & NNenum? \\\hline
	MNIST 6$\times$100 & 99\% & 90\% & 26\% & 30\% & ?\%   \\
	$\epsilon = 0.026$ & &  & 6520s & 7000s &  ?s  \\  \hline
\end{tabular}
\caption{Complete verifiers run on the smallest of the considered DNN. 
Because the instances are hard, most test time-out at the global threshold of 10.000s per image, and no complete verifier could outperform 
$\alpha,\beta$-CROWN (40\%) for this DNN, despite the much larger average runtime. We didnt test on larger DNNs as this small one already took considerable amount of time.}
\label{table_complete}
\end{table}

@@

Gap between verified and attackable images. Lower is better.
M for MNIST, C for CIFAR
%The table of undecided images among first 100 images in database by different methods for different network.

\begin{table}
\centering
\begin{tabular}{||l||c|c|c|c||c||}
	\hline \hline
	 & $\alpha,\beta$-Crown & $\alpha,\beta$-Crown & $\alpha,\beta$-Crown & Refined & Hybrid \\ 
	 Network & TO=10s & TO=30s & TO=2000s & $\beta$-Crown & MILP \\ 
	\hline
	MNIST 6$\times$100 & 57\% (6.9s) & 55\% (18.9s) & 50\% (1026s) & 13\% 
	(92s) & 13\% \bf (46s) \\ \hline
	MNIST 6$\times$200 & 50\% (6.5s) & 47\% (17s) & 46\% (930s) & 9\% (80s) & \bf 8\% (71s) \\ \hline
	MNIST 9$\times$100 & 63\% (7.2s) & 58\% (20s) & 58\% (1163s) & 21\% (102s) & \bf 15\% (61s) \\ \hline
	MNIST 9$\times$200 & 56\% (6.8s) & 55\% (18s) & 54\% (1083s) & 16\% (83s) & \bf 8\% (78s) \\ \hline
	MNIST 6$\times$500 & 53\% (6.4s) & 51\% (16s) & 50\% (1002s) & N/A & 
	\bf 10\% (402s) \\ \hline
	CIFAR CNN-B-adv & 28\% (4.3s) & 22\% (8.7s) & 20\% (373s) & N/A & \bf 11\% (417s) \\ \hline \hline
	CIFAR Resnet & 0\% (2s) & 0\% (2s) & 0\% (2s) & N/A & 0\% (2s) \\ \hline \hline
\end{tabular}
\caption{Gap $\%$ between attackable images and verifiable images ({\em lower is better}), as certified by $\alpha,\beta$-Crown, refined $\beta$-Crown and Hybrid MILP, run on 7 DNNs (average runtime per image in second). The 6 first DNNs are hard instances, while the last one (ResNet) is an easy instance (it is trained to be easy to verify using Wong, but with a very low accuracy level), provided for reference.}
\label{table_hybrid}
\end{table}

### 2024-09-16 ICLR25/tables.tex
update tables.tex
table updated in main document

\begin{table}
	\centering
	\caption{$\alpha,\beta$-Crown with pure BaB}
	\label{table_beta}
	\begin{tabular}{||c|c|c||c|c|c||}
		\hline
		Network & Acc & Upper  & $\alpha,\beta$-Crown& $\alpha,\beta$-Crown & $\alpha,\beta$-Crown \\ 
		Perturbation &   & Bound & TO=10s & TO=30s & TO=2000s\\ \hline

		MNIST 6$\times$100 & 99\% & 90\% & 33\% & 35\% & 40\%   \\
		$\epsilon = 0.026$ &  &  & 6.9s &  18.9s &  1026.1s  \\  \hline
		MNIST 6$\times$200 & 99\%  & 96\%  & 46\%  & 49\%  & 50\%   \\ 
		$\epsilon = 0.015$ & &  & 6.5s &  16.6s &  930.3s  \\  \hline
		MNIST 9$\times$100 & 97\%  & 86\%  & 23\%  & 28\%  & 28\%   \\
		$\epsilon = 0.026$ &  &  & 7.2s &  20.1s &  930.3s  \\  \hline
		MNIST 9$\times$200 & 97\%  & 91\%  & 35\%  & 36\%  & 37\%   \\ 
		$\epsilon = 0.015$ & &  & 6.8s &  18.2s &  1083.1s  \\  \hline
		MNIST 6$\times$500 & 100\%  & 94\%  & 41\%  & 43\%  & 44\%   \\ 
		$\epsilon = 0.035$ & &  & 6.4s &  16.4s &  1002.5s  \\  \hline
		CIFAR CNN-B-adv & 78\%  & 62\%  &  34\% & 40\%  & 42\%   \\
		$\epsilon = 2/255$&  &  & 4.3s & 8.7s & 373s  \\ \hline \hline
		CIFAR ResNet & 29\%  & 25\%  & 25\%  & 25\%  & 25\%   \\
		$\epsilon = 2/255$ &  &  & 2s & 2s & 2s  \\ \hline
	\end{tabular}
\end{table}

@@

\begin{table}
	\centering
	\caption{SAME FORMAT for complete verifiers!!!}
	\label{table_complete}
	\begin{tabular}{||c||c|c|c||}
		\hline
		Network &  Marabou  & Full MILP & NNenum? \\ \hline
		MNIST 6$\times$100 & 26\% & 30\% & ?\%   \\ \hline
		$\epsilon = 0.026$ & 6520s & 7000s
		&  ?s  \\  \hline
	\end{tabular}
\end{table}

>>>>

\begin{table}
	\centering
	\begin{tabular}{||c|c|c||c|c|c||}
		\hline
		Network & Acc & Upper  & $\alpha,\beta$-Crown& $\alpha,\beta$-Crown & $\alpha,\beta$-Crown \\ 
		Perturbation &   & Bound & TO=10s & TO=30s & TO=2000s\\ \hline
		MNIST 6$\times$100 & 99\% & 90\% & 33\% & 35\% & 40\%   \\
		$\epsilon = 0.026$ &  &  & 6.9s &  18.9s &  1026.1s  \\  \hline
		MNIST 6$\times$200 & 99\%  & 96\%  & 46\%  & 49\%  & 50\%   \\ 
		$\epsilon = 0.015$ & &  & 6.5s &  16.6s &  930.3s  \\  \hline
		MNIST 9$\times$100 & 97\%  & 86\%  & 23\%  & 28\%  & 28\%   \\
		$\epsilon = 0.026$ &  &  & 7.2s &  20.1s &  930.3s  \\  \hline
		MNIST 9$\times$200 & 97\%  & 91\%  & 35\%  & 36\%  & 37\%   \\ 
		$\epsilon = 0.015$ & &  & 6.8s &  18.2s &  1083.1s  \\  \hline
		MNIST 6$\times$500 & 100\%  & 94\%  & 41\%  & 43\%  & 44\%   \\ 
		$\epsilon = 0.035$ & &  & 6.4s &  16.4s &  1002.5s  \\  \hline
		CIFAR CNN-B-adv & 78\%  & 62\%  &  34\% & 40\%  & 42\%   \\
		$\epsilon = 2/255$&  &  & 4.3s & 8.7s & 373s  \\ \hline \hline
		CIFAR ResNet & 29\%  & 25\%  & 25\%  & 25\%  & 25\%   \\
		$\epsilon = 2/255$ &  &  & 2s & 2s & 2s  \\ \hline
	\end{tabular}
	\caption{$\alpha,\beta$-Crown with pure BaB. 6 first Networks are hard instances, while the last one (ResNet) is an easy instance (it is trained to be easy to verify using Wong, but with a very low accuracy level), provided for reference.}
	\label{table_beta}
\end{table}

@@

\begin{table}
	\centering
	\caption{SAME FORMAT for complete verifiers!!!}
	\label{table_complete}
	\begin{tabular}{||c||c|c|c||}
		\hline
		Network &  Marabou  & Full MILP & NNenum? \\ \hline
		MNIST 6$\times$100 & 28\% & 39\% & ?\%   \\ \hline
		$\epsilon = 0.026$ & 6200s & 7000s
		&  ?s  \\  \hline
	\end{tabular}
\end{table}
